Abstract Binding Trees

### 2015-10-25 paper.tex
add refinements for variables & metavariables

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\OMode{#5}%
}
\newcommand\IsWf[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}\in_{\mathbf{wf}}\OMode{#5}%
}

@@

Having specified an approximation of the grammar of telescopes and sequents in
the abt logical framework, we can proceed to define proper wellformedness via
\emph{inductive refinement}. The basic idea is to introduce a new form of
(meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{s}$ which expresses the
extrinsic wellformedness properties we wish to verify, presupposing
$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$.

%% TODO: implement refinement judgment for wellscopedness

For the sake of this example, we will just stipulate that anything of sort $\SortExp$ or $\SortType$ is
grammatical:
\begin{align*}
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{A}{\SortType}
  }{
  }\qquad
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\SortExp}
  }{
  }
\end{align*}

>>>>

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\OMode{#5}%
}
\newcommand\IsWf[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
}

@@

Having specified an approximation of the grammar of telescopes and sequents in
the abt logical framework, we can proceed to define proper wellformedness via
\emph{inductive refinement}. The basic idea is to introduce a new form of
(meta)-judgment $\IsWf{\Omega}{\Upsilon}{\Gamma}{M}{s}$ which expresses the
extrinsic wellformedness properties we wish to verify, presupposing
$\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$. Additionally, we introduce an
analogous judgment on bound terms, $\IsWf{\Omega}{\Upsilon}{\Gamma}{E}{v}$
presupposing $\IsAbt{\Omega}{\Upsilon}{\Gamma}{E}{v}$, defined uniformly as
follows:
\[
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkValence{\vec{p}}{\vec{q}}{s}
    }
  }{
    \IsWf{\Omega}{\Upsilon,\vec{u}:\vec{p}}{\Gamma,\vec{x}:\vec{q}}{M}{s}
  }
\]

Likewise, wellformedness for variables and metavariables is defined uniformly:
\begin{gather*}
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{x}{s}
  }{
  }\qquad
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{
      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
    }{s}
  }{
    \begin{array}{l}
      \Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{q_0,\dots,q_n}{s}}\\
      \HypJ{
        \IsWf{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }
\end{gather*}

\begin{remark}
  Note that the refinement for variables $x$ is not trivial, since it is only
  defined in case the presupposition $\IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}$
  is satisfied.
\end{remark}

The remainder of the definition of refinement proceeds by induction on sorts
and operators. For the sake of this example, we will just stipulate that
anything of sort $\SortExp$ or $\SortType$ is grammatical:

\begin{align*}
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{A}{\SortType}
  }{
  }\qquad
  \infer{
    \IsWf{\Omega}{\Upsilon}{\Gamma}{M}{\SortExp}
  }{
  }
\end{align*}
